---
title: "data import/manipulation"
output: github_document
---
# Section1: data import 

```{r setup}
library(tidyverse)
```

## Read in some data

read in the litters dataset

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
 "." means the current folder. This is a relative path. (absolute path not recommended).
 
 "janitor::" means we want to use the command in this library. we can also load janitor in library ().
 
 
## take a look at the data

printing
```{r}
litters_df
```

the first and last few columns of dataset.
```{r}
head(litters_df)
tail(litters_df)
```


the no. of missing values for char and num vars and descriptive stats for num, and small histogram.
```{r}
skimr::skim(litters_df)
```

view; (can put in console.)
```{r}
view(litters_df)
```


##options to read csv
skip the first 10 rows. do not include column names. 
```{r}
litters_df = read_csv("./data/FAS_litters.csv",skip = 10,col_names  = FALSE, 
                      na = c(".", "", "NA","999"))
```

check out `?read_csv()` for more information and other options.

## other file formats

#####Read in an excel file

```{r}
library(readxl)
```
This could be put at top of the files.



```{r}
mlb_df = read_excel("./data/mlb11.xlsx",range = "A1:F7")
mlb_df
```
can use `view(mlb_df)` in the console to see the entire dataset

check out `?read_excel` for options.

##### read in a SAS file

```{r}
library(haven)
```

```{r}
pulse_df = read_sas("./data/public_pulse_data.sas7bdat")
pulse_df
```

## Comparison with base R

What about `read.csv`
```{r}
litters_readr = read_csv("./data/FAS_litters.csv")
litters.base = read_csv("./data/FAS_litters.csv")

litters.base
litters_readr
```

better to use read_csv.


## exporting data

export the mlb sub table
```{r}
write_csv(mlb_df,"./data/mlb_subtable.csv")
```


# Section 2: data manipulation

## `select`

choose some columns but not the others.

```{r}
select(litters_df, group, litter_number)
```

can select groups of var
```{r}
select(litters_df, gd0_weight:gd_of_birth)
```

get rid of a column
```{r}
select(litters_df, -litter_number)
```

renaming columns and select at the same time
```{r}
select(litters_df, GROUP = group, LITTerNumber = litter_number)
```

rename columns
```{r}
rename(litters_df, GROUP = group, LITTerNumber = litter_number)
```


select helpers. use `?select_helpers`

```{r}
select(litters_df, starts_with("gd"))
```

put litters_number as the 1st column, and keep all other columns
```{r}
select(litters_df, litters_number, everything())
```
=
```{r}
relocate(litters_df, litters_number)
```


## `filter`

operate on rows. need to give conditions to it

```{r}
filter(litters_df, gd0_weight < 22)
```

== is a test for equality (if =22 is true)
```{r}
filter(litters_df, gd0_weight == 22)
```

! the condition in the bracket is not true
```{r}
filter(litters_df, !(gd0_weight == 22))
```

```{r}
filter(litters_df, !(gd0_weight == 22))
```

```{r}
filter(litters_df, group %in% c("Mod8", "Con"))
```