---
title: "data import"
output: github_document
---

```{r setup}
library(tidyverse)
```

## Read in some data

read in the litters dataset

```{r}
litters_df = read_csv("./data/FAS_litters.csv")
litters_df = janitor::clean_names(litters_df)
```
 "." means the current folder. This is a relative path. (absolute path not recommended).
 
 "janitor::" means we want to use the command in this library. we can also load janitor in library ().
 
 
## take a look at the data

printing
```{r}
litters_df
```

the first and last few columns of dataset.
```{r}
head(litters_df)
tail(litters_df)
```


the no. of missing values for char and num vars and descriptive stats for num, and small histogram.
```{r}
skimr::skim(litters_df)
```

view; (can put in console.)
```{r}
view(litters_df)
```


##options to read csv
skip the first 10 rows. do not include column names. 
```{r}
litters_df = read_csv("./data/FAS_litters.csv",skip = 10,col_names  = FALSE, 
                      na = c(".", "", "NA","999"))
```

check out `?read_csv()` for more information and other options.
